---
title: "Model2:LogOdds"
output: html_document
---

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyr)
library(data.table)
library(tibble)
library(mltools)
library(aod)
```

```{r}
chess_games_analyzed = read.csv('../Data/chess_games_analyzed.csv')
#run last 3 code blocks to get chess_games_pieces_remaining
chess_games_pieces_remaining = read.csv('../Data/chess_games_pieces_remaining.csv')
```

```{r}
outcome = as.factor(chess_games_pieces_remaining$winner)
one_hot_winner = one_hot(data.table('outcome' = outcome))

chess_games_pieces_remaining[,14:16] = one_hot_winner
chess_games_pieces_remaining
```

```{r}
whitelogit <- glm(outcome_white ~ white_bishop_rem + white_knight_rem + white_pawn_rem +
                 white_queen_rem + white_rook_rem, data = chess_games_pieces_remaining)
confint(whitelogit)
```

```{r}
blacklogit <- glm(outcome_black ~ black_bishop_rem + black_knight_rem + black_pawn_rem +
                 black_queen_rem + black_rook_rem, data = chess_games_pieces_remaining)
confint(blacklogit)
```

Inspired by Rasmus Bååth at R Bloggers - look at remaining piece difference

https://www.r-bloggers.com/2015/06/big-data-and-chess-what-are-the-predictive-point-values-of-chess-pieces/
```{r}
chess_games_pieces_remaining = chess_games_pieces_remaining %>%
  mutate(pawn_diff = white_pawn_rem - black_pawn_rem) %>%
  mutate(knight_diff = white_knight_rem - black_knight_rem) %>%
  mutate(bishop_diff = white_bishop_rem - black_bishop_rem) %>%
  mutate(rook_diff = white_rook_rem - black_rook_rem) %>%
  mutate(queen_diff = white_queen_rem - black_queen_rem)
```

Never used but could be valuable:
```{r}
logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}
```


```{r}
alllogit = glm(outcome_white ~ 
                 1 + pawn_diff + knight_diff + bishop_diff + rook_diff + queen_diff,
               data = chess_games_pieces_remaining, family = 'binomial')
summary(alllogit)
confint(alllogit)

coef_df = data.frame(logodds = coefficients(alllogit))*-1
coef_df <- tibble::rownames_to_column(coef_df, "Piece")
coef_df$Piece <- factor(coef_df$Piece, levels = coef_df$Piece)

coef_df
```

```{r}
ggplot(coef_df, aes(Piece, logodds)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = coef_df,
            aes(x = Piece, y = logodds, 
                label=round(logodds,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 1.5) +
  labs(
    title = 'Log Odds Value of Pieces',
    subtitle = 'Chess Piece proposed relative value based on logistic regression model',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_LogOddsOverallValue.pdf")
```

Numbers are hard to interpret here, lets scale by setting the pawn's logodds to 1

```{r}
pawn_logodds = coef_df[2,]$logodds
pawnscalefactor = 1 / pawn_logodds

coef_df$logodds_pawnscaled = coef_df$logodds * pawnscalefactor
```

```{r}
knight_logodds = coef_df[3,]$logodds
knightscalefactor = 3 / knight_logodds

coef_df$logodds_knightscaled = coef_df$logodds * knightscalefactor
coef_df
```
```{r}
ggplot(coef_df, aes(Piece, logodds_knightscaled)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = coef_df,
            aes(x = Piece, y = logodds_knightscaled, 
                label=round(logodds_knightscaled,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 1.5) +
  labs(
    title = 'Log Odds Value of Pieces, Scaled to Knight = 3',
    subtitle = 'Chess Piece proposed relative value based on log regression model, with knight set at 3.0',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_LogOddsOverallValue_KnightScaled.pdf")
```

Repeat, but just for Novice and Class A
```{r}
novice_data_remaining = 
  chess_games_pieces_remaining[chess_games_pieces_remaining$rating_category == "Novice",]
classd_data_remaining = 
  chess_games_pieces_remaining[chess_games_pieces_remaining$rating_category == "ClassD",]
classc_data_remaining = 
  chess_games_pieces_remaining[chess_games_pieces_remaining$rating_category == "ClassC",]
classb_data_remaining = 
  chess_games_pieces_remaining[chess_games_pieces_remaining$rating_category == "ClassB",]
classa_data_remaining = 
  chess_games_pieces_remaining[chess_games_pieces_remaining$rating_category == "ClassA",]
```

In full transparency, should have definetly functionilzed this, but was running a bit short on time, and did not have time for a potential debugging.


Novice

```{r}
novice_alllogit = glm(outcome_white ~ 
                 1 + pawn_diff + knight_diff + bishop_diff + rook_diff + queen_diff,
               data = novice_data_remaining, family = 'binomial')

#summary(novice_alllogit)
#confint(novice_alllogit)

novice_coef_df = data.frame(logodds = coefficients(novice_alllogit))*-1
novice_coef_df <- tibble::rownames_to_column(novice_coef_df, "Piece")
novice_coef_df$Piece <- factor(novice_coef_df$Piece, levels = novice_coef_df$Piece)

novice_knight_logodds = novice_coef_df[3,]$logodds
novice_knightscalefactor = 3 / novice_knight_logodds

novice_coef_df$logodds_knightscaled = novice_coef_df$logodds * novice_knightscalefactor
novice_coef_df
```

```{r}
ggplot(novice_coef_df, aes(Piece, logodds_knightscaled)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = novice_coef_df,
            aes(x = Piece, y = logodds_knightscaled, 
                label=round(logodds_knightscaled,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 2.0) +
  labs(
    title = 'Log Odds Value of Novice Pieces, Scaled to Knight = 3',
    subtitle = 'Novice Chess Pieces proposed relative value based on log reg model, with knight set at 3.0',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_NoviceLogOddsValue_KnightScaled.pdf")
```

Class D

```{r}
classd_alllogit = glm(outcome_white ~ 
                 1 + pawn_diff + knight_diff + bishop_diff + rook_diff + queen_diff,
               data = classd_data_remaining, family = 'binomial')

#summary(novice_alllogit)
#confint(novice_alllogit)

classd_coef_df = data.frame(logodds = coefficients(classd_alllogit))*-1
classd_coef_df <- tibble::rownames_to_column(classd_coef_df, "Piece")
classd_coef_df$Piece <- factor(classd_coef_df$Piece, levels = novice_coef_df$Piece)

classd_knight_logodds = classd_coef_df[3,]$logodds
classd_knightscalefactor = 3 / classd_knight_logodds

classd_coef_df$logodds_knightscaled = classd_coef_df$logodds * classd_knightscalefactor
classd_coef_df
```

```{r}
ggplot(classd_coef_df, aes(Piece, logodds_knightscaled)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = classd_coef_df,
            aes(x = Piece, y = logodds_knightscaled, 
                label=round(logodds_knightscaled,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 2.0) +
  labs(
    title = 'Log Odds Value of Class D Pieces, Scaled to Knight = 3',
    subtitle = 'Class D Chess Pieces proposed relative value based on log reg model, with knight set at 3.0',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_ClassDLogOddsValue_KnightScaled.pdf")
```

Class C

```{r}
classc_alllogit = glm(outcome_white ~ 
                 1 + pawn_diff + knight_diff + bishop_diff + rook_diff + queen_diff,
               data = classc_data_remaining, family = 'binomial')

#summary(novice_alllogit)
#confint(novice_alllogit)

classc_coef_df = data.frame(logodds = coefficients(classc_alllogit))*-1
classc_coef_df <- tibble::rownames_to_column(classc_coef_df, "Piece")
classc_coef_df$Piece <- factor(classc_coef_df$Piece, levels = classc_coef_df$Piece)

classc_knight_logodds = classc_coef_df[3,]$logodds
classc_knightscalefactor = 3 / classc_knight_logodds

classc_coef_df$logodds_knightscaled = classc_coef_df$logodds * classc_knightscalefactor
classc_coef_df
```

```{r}
ggplot(classc_coef_df, aes(Piece, logodds_knightscaled)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = classc_coef_df,
            aes(x = Piece, y = logodds_knightscaled, 
                label=round(logodds_knightscaled,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 2.0) +
  labs(
    title = 'Log Odds Value of Class C Pieces, Scaled to Knight = 3',
    subtitle = 'Class C Chess Pieces proposed relative value based on log reg model, with knight set at 3.0',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_ClassCLogOddsValue_KnightScaled.pdf")
```


Class B

```{r}
classb_alllogit = glm(outcome_white ~ 
                 1 + pawn_diff + knight_diff + bishop_diff + rook_diff + queen_diff,
               data = classb_data_remaining, family = 'binomial')

#summary(novice_alllogit)
#confint(novice_alllogit)

classb_coef_df = data.frame(logodds = coefficients(classb_alllogit))*-1
classb_coef_df <- tibble::rownames_to_column(classb_coef_df, "Piece")
classb_coef_df$Piece <- factor(classb_coef_df$Piece, levels = classb_coef_df$Piece)

classb_knight_logodds = classb_coef_df[3,]$logodds
classb_knightscalefactor = 3 / classb_knight_logodds

classb_coef_df$logodds_knightscaled = classb_coef_df$logodds * classb_knightscalefactor
classb_coef_df
```

```{r}
ggplot(classb_coef_df, aes(Piece, logodds_knightscaled)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = classb_coef_df,
            aes(x = Piece, y = logodds_knightscaled, 
                label=round(logodds_knightscaled,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 2.0) +
  labs(
    title = 'Log Odds Value of Class B Pieces, Scaled to Knight = 3',
    subtitle = 'Class B Chess Pieces proposed relative value based on log reg model, with knight set at 3.0',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_ClassBLogOddsValue_KnightScaled.pdf")
```

```{r}
classa_alllogit = glm(outcome_white ~ 
                 1 + pawn_diff + knight_diff + bishop_diff + rook_diff + queen_diff,
               data = classa_data_remaining, family = 'binomial')

#summary(novice_alllogit)
#confint(novice_alllogit)

classa_coef_df = data.frame(logodds = coefficients(classa_alllogit))*-1
classa_coef_df <- tibble::rownames_to_column(classa_coef_df, "Piece")
classa_coef_df$Piece <- factor(classa_coef_df$Piece, levels = classa_coef_df$Piece)

classa_knight_logodds = classa_coef_df[3,]$logodds
classa_knightscalefactor = 3 / classa_knight_logodds

classa_coef_df$logodds_knightscaled = classa_coef_df$logodds * classa_knightscalefactor
classa_coef_df
```

```{r}
ggplot(classa_coef_df, aes(Piece, logodds_knightscaled)) + 
  geom_col(fill = 'dark green') +
  coord_flip() +
  geom_text(data = classa_coef_df,
            aes(x = Piece, y = logodds_knightscaled, 
                label=round(logodds_knightscaled,2)),
            color="white", size=3.5,
            position = position_dodge(width = .9),
            hjust = 2.0) +
  labs(
    title = 'Log Odds Value of Class A Pieces, Scaled to Knight = 3',
    subtitle = 'Class A Chess Pieces proposed relative value based on log reg model, with knight set at 3.0',
    y = "LogOdds Value", 
    x = "Piece Difference"
  )

ggsave("../Images/Model2_ClassALogOddsValue_KnightScaled.pdf")
```

```{r}
compare_coef_df = novice_coef_df
compare_coef_df$Novice_logodds = compare_coef_df$logodds_knightscaled
compare_coef_df$logodds_knightscaled = NA
compare_coef_df$ClassA_logodds = classa_coef_df$logodds_knightscaled
compare_coef_df = compare_coef_df[,c(1,4,5)]
compare_coef_df

dfm <- melt(compare_coef_df[,colnames(compare_coef_df)],
            id.vars = 1)

colnames(dfm) <- c('Piece', 'Rating Category', 'LogOdds')

ggplot(dfm, aes(fill = `Rating Category`, y = LogOdds, x = Piece)) + 
  geom_bar(position="dodge", stat="identity") + 
  geom_text(data = dfm,
            aes(x = Piece, y = LogOdds, 
                label=round(LogOdds, 2),
                group = `Rating Category`),
            color="black", size=3.5,
            position = position_dodge(width = .9),
            hjust = -0.1) +
  labs(
    title = 'Log Odds Value, Scaled to Knight = 3',
    subtitle = 'Comparing Class A and Novice Logistic Regression Values',
    x = 'Piece Difference',
    y = 'Log Odds Value, Knight Scaled'
  ) + 
  coord_flip()

```


```{r}
compare_coef_df_full = data.frame(compare_coef_df)
compare_coef_df_full$ClassB_logodds = classb_coef_df$logodds_knightscaled
compare_coef_df_full$ClassC_logodds = classc_coef_df$logodds_knightscaled
compare_coef_df_full$ClassD_logodds = classd_coef_df$logodds_knightscaled
compare_coef_df_full = compare_coef_df_full[, c(1,2,6,5,4,3)]
compare_coef_df_full[,c(2:6)] = round(compare_coef_df_full[,c(2:6)], 2)
compare_coef_df_full

library(gridExtra)
pdf("../Images/Model2_Valuation.pdf", height=11, width=8.5)
grid.table(compare_coef_df_full)
dev.off()

```





Combine lists of all pieces taken. 
Make sure not to run this more than once
```{r}
#chess_games_analyzed_2 = chess_games_analyzed %>%
#  unite('white_allpieces_taken', 41:46, remove = FALSE) %>%
#  unite('black_allpieces_taken', 72:77, remove = FALSE)
```



Create function to find what pieces are remaining
```{r}
#endofgame_pieces = function(lst_allpiecestaken){
#  
#  #clean list of all pieces taken and turn it into a vector
#  cleanlst_allpiecestaken = str_replace_all(lst_allpiecestaken, "_", "")
#  cleanlst_allpiecestaken = str_split(cleanlst_allpiecestaken, " ")[[1]]
#  cleanlst_allpiecestaken = cleanlst_allpiecestaken[nzchar(cleanlst_allpiecestaken)]
#  
#  #create summarized table
#  df = data.frame(table(cleanlst_allpiecestaken))
#  
#  #EDGE CASE
#  if(length(df) == 1){
#    df = data.frame(cleanlst_allpiecestaken = 'Pawn',
#           Freq = 0)
#  }
#  
#  #clean dataframe
#  df$cleanlst_allpiecestaken = as.character(df$cleanlst_allpiecestaken)
#  
#  #fill in such that we have each piece in the dataframe 
#  if(!("Pawn" %in% df$cleanlst_allpiecestaken)) {
#    df[nrow(df) + 1,] = c('Pawn', 0)
#  }
#  if(!("Knight" %in% df$cleanlst_allpiecestaken)) {
#    df[nrow(df) + 1,] = c('Knight', 0)
#  }
#  if(!("Bishop" %in% df$cleanlst_allpiecestaken)) {
#    df[nrow(df) + 1,] = c('Bishop', 0)
#  }
#  if(!("Rook" %in% df$cleanlst_allpiecestaken)) {
#    df[nrow(df) + 1,] = c('Rook', 0)
#  }
#  if(!("Queen" %in% df$cleanlst_allpiecestaken)) {
#    df[nrow(df) + 1,] = c('Queen', 0)
#  }
#  
#  df = df %>%
#    #arrange dataframe alphabetically so intial_pieces appending is clean
#    arrange(cleanlst_allpiecestaken) %>%
#    mutate(initial_pieces = c(2,2,8,1,2)) %>%
#    mutate(Freq = as.numeric(Freq)) %>%
#    mutate(initial_pieces = as.numeric(initial_pieces)) %>%
#    #find remaining pieces for each game 
#    mutate(remaining_pieces = initial_pieces - Freq)
#  
#  #EDGE: for promotions, there may be more than the initial pieces, leading to a -1 in the #remaining pieces column. Therefore, make every negative value a positive one
#  
#  df[,4][df[,4] < 0] <- 0
#  
#  
#  #transpose dataframe to be able to append to dataframe
#  df = df[,c(1,4)]
#  df_t = transpose(df)
#  colnames(df_t) = paste0(df_t[1,], "_remaining")
#  df_t = df_t[2,]
#  
#  
#  return(df_t)
#}

```


All rows
```{r}
#endofgame_pieces(chess_games_analyzed_2$white_allpieces_taken[2])[1,c(1:5)]

#chess_games_analyzed_2 = chess_games_analyzed_2 %>%
#  add_column(white_bishop_rem = NA,
#             white_knight_rem = NA,
#             white_pawn_rem = NA,
#             white_queen_rem = NA,
#             white_rook_rem = NA,
#             black_bishop_rem = NA,
#             black_knight_rem = NA,
#             black_pawn_rem = NA,
#             black_queen_rem = NA,
#             black_rook_rem = NA)
  

#for(i in c(1:nrow(chess_games_analyzed_2))) {
  
#  print(i)
  
#  white_pieces = chess_games_analyzed_2$white_allpieces_taken[i]
#  black_pieces = chess_games_analyzed_2$black_allpieces_taken[i]
  
#  print('step 1')
  
#  white_df = endofgame_pieces(white_pieces)
#  black_df = endofgame_pieces(black_pieces)
  
#  print('step 2')
  
  
#  chess_games_analyzed_2[i,84:88] = white_df[1,c(1:5)]
#  chess_games_analyzed_2[i,89:93] = black_df[1,c(1:5)]
  
#}

#Subset dataframe to only include relevant columns

#chess_games_pieces_remaining = chess_games_analyzed_2[,c(7,82:93)]

#write.csv(chess_games_pieces_remaining,"../Data/chess_games_pieces_remaining.csv", row.names = FALSE)
```



